### 2019-10-09 peer-review/review-response.Rmd

*Thank you for your helpful comments on this paper.*

### Introduction

1.	Page 4, Line 21: BCG efficacy has been known with great heterogeneity and it affects the assessment of vaccine impact on TB burden. It is therefore commonly considered in decision-making for BCG vaccination policy. I suggest that the authors include the uncertainty of BCG efficacy in this analysis and compared the relative importance to the assumption of how TB burden declines over time.

*We agree with the reviewer that this is an important issue. We have included some of this uncertainty by re-estimating incidence rates for both those BCG vaccinated and unvaccinated. These rates are then used to estimate vaccine efficacy. This approach makes use of the available data on the effiacy of the BCG vaccine in England and Wales. Including additional uncertainty observed globally would be inappropriate here as there is little evidence for this amongst the White UK born population. This does mean that the findings from this model reflect the impact in these countries only and cannot be directly generalised.*

@@

*Thank you. We have corrected this throughout*

### Methods

3.	Page 5, Line 6-7: Is the ‘transmission chain model’ included in the Sutherland et al.’s model or developed by the authors? Please state it here.

*We have extensively reworded and extended this section (please see the marked up version). As part of ths we have added the following sentence for clarity,*

@@

4.	Page 5, steps for defining the transmission model: Presenting equations of the model is certainly helpful for understanding the estimation methods. In addition to the descriptions of equation components, I encourage the authors to explain more about the rationales behind. For example, how to interpret ‘initial generation size (x)’? What is the purpose of relating step 1 to step 2? This kind of additional information will allow readers from a wide range of background to understand the model easily.

*Whilst this information was present in the original Sutherland et al. paper we agree that more detail was needed here. We have therefore added greater explanatory detail for each step (please see the marked up paper).*

5.	Page 5, last paragraph: Is the ‘distribution of secondary cases (N)’ estimated by the three age groups (of 15-19, 20-24, and 25-29 year old)? The power term ‘5-z’ of the decay percentage (d) needs additional explanations in the text.

*Yes the distribution of cases is estimated by age group. We have reformulated these equations for greater clarity (please see the marked up paper). Note that due to paper changes this section of the model has been moved to the SI.*

@@

*Yes, thank you - corrected.*

8.	Page 6, Line 28-32: This part is too long to follow and with sentence fragments. Please re-organise it.

*Reworded to:*

*"Data collection for the ETS began in 2000 and prior to this notification data was only available in years with notifications surveys (1973, 1978, and 1983). We therefore estimated incidence rates between 1984 and 1999, and for the years between notifications surveys (1974-1977 and 1979-1982), using locally estimated scatterplot smoothing (LEOSS) regression fitted to incidence rates published in [11] and the estimated incidence rates from 2000 on-wards. LOESS is a local regression method that combines multiple regression models in a k-nearest neighbours meta-model.[@FoxSnow:2019] This approach allows nonlinear trends to be fitted using a series of linear models. For years prior to 1973 the annual decreases were assumed to be the mean of the annual decreases from the previous 3 years. For both proxy measures the annual decreases in incidence rates post 2016 were assumed to be the average of the estimates in 2013, 2014, and 2015."*

@@

*We agree that this needed further explanation and that the definition used in 1. of Sutherland et al. was incorrect. We have corrected the Sutherland model and extended the model explanation to include this.*

I also found it harder than it should have been to understand exactly what was different between this analysis and the original analysis, and which differences mattered most (which is important as it is the main aim of the work as I understand it). I wonder if a table would be useful detailing the differences: any implementation/modelling differences; differences in point estimates of d (both from methods and use of more recent data?); differences in parameters from literature; inclusion of uncertainty in each. I think it's important to differentiate between differences that could have been included in the analysis at the time (eg uncertainty) vs things that have emerged since (eg updated evidence on serial intervals). I also think it's important to be clear whether methods, central estimates, or uncertainty matter most from the point of view of decision-relevant changes in the evidence generated.

*We have introduced a subsection of the model explanation detailing our additions to the model. The Updating model parameter estimates and incorporating parameter uncertainty section of the methods contains a discussion of differences between the original and updated analysis. We have reworded this for clarity (please see the marked up paper). As you say, parameter uncertainty for incidence rates could have been included at the time but was not, whilst the updated value for the serial interval has only recently become available. This is now explicitly stated in the reworded methods paragraph.*

*We agree that it is important to be clear which findings are relevant to policy makers and in what context. We have reworded the 4th discussion paragraph to provide more context.*

@@

*Thank you - corrected throughout.*

- In a couple of places (ie strengths/limitations bullets & Discussion) it is stated this is not "the most accurate method" for assessing the impact of these changes. This does beg questions about what other methods would have been better, and in what respects (eg how is "accuracy" defined?). This is worth discussing. I think it also means that the rationale for the work is about assessing the imapct of updated/modified analysis for inputs within an existing framework, rather than trying to provide a better answer to the same question (otherwise, why use a suboptimal method?)

*We agree that this could have been better phrased. We have rephrased as follows,*

*"The simulation approach used here, although updated where possible, is not the most accurate method for estimating the impact of ending the BCG schools scheme as it relies on numerous assumptions based on the available knowledge in 1987 and does not account for the role of non-White and non-UK born cases. However, the strength of this work is that the estimates are based on the framework used to inform policy making. This allowed the strength of the model used in the decision-making process to be assessed once parameter uncertainty had been incorporated and for flaws in the model to be identified. This would not have been possible if the impact had been assessed using only the observed data or with a alternative model. It also allowed estimates based on updated data to be compared to historic estimates within the same framework. This would also not have been possible if a different framework had been used."*

@@

- "TB notification surveys" - could you explain what these are please?

*Added detail to the methods. Please see the marked up paper.*

- ref 17: I thought there had been a systematic review of serial intervals for TB (Ma et al, Epidemiology & Infection). Justify why using this study in particular please. Choice of evidence an important part of paper since arguing to update & compare.

*Thanks for pointing this out. Whilst there has been a systematic review the authors identifed only two papers with estimates (the one used here and one other older paper). Both papers made use of similar datasets in the Netherlands, however, the cited paper is most recent (by 10 years) and has an improved sample size and improved methods compared to the older paper.*

@@

- pg 6 paragraph on methods to quantify decline: again, I found this confusing and it is a key part of the analysis. Also, what is "local regression" please?

*We have reworded this paragraph (and made additional changes suggested by reviewer 1). Please see the marked up paper. We have also included a graph in the results from the SI that shows  the changes over time in the annual decline depending on the data/assumptions used.*

@@

*We have added the following clarification for local regression,*

*"LOESS (locally estimated scatterplot smoothing) is a local regression method. It combines multiple regression models in a k-nearest neighbours meta-model. This approach allows nonlinear trends to be fitted using a series of linear models."*

@@

*We agree that these estimates (from the original Sutherland et al. paper) are very unrealistic. Sutherland et al. assumed a year-on-year decline of 9% with an estimated incidence of  ~ 23 per 100,000 in 1983 for the unvaccinated population. In 2013 this would give an incidence rate of ~ 2.6. These very small incidence rates mean that the difference between incidence rates in unvaccinated and vaccinated cases is also very small leading to the spuriously high number need to vaccinate. In comparison a 2% decline would give an inidence rate of ~ 12.5. Our additions to the methods section add extra detail on how these estimates are arrived at.*

- I think discounting needs to be discussed at some point. This is almost always included in decision analyses.

*We agree that discounting should be included for economic evaluation. However we have not seen it done widely when the study is focusing purely on the epidemiological impact of vaccination. We have added the following sentences to the discussion,*

*"Policy makers should consider these updated estimates when assessing the role of BCG vaccination in those at school-age. However, decisions regarding vaccine policy in the UK require economic evaluation, which discounts costs and benefits in the future; discounting has not been applied in this study which estimates the epidemiological impact of vaccination only."*

### 2019-10-10 peer-review/review-response.Rmd

*We have revised the patient and public involvement statement to the following,*

*"We did not involve patients or the public in the design or planning of this study."*

- Please complete and include a TRIPOD checklist, ensuring that all points are included and state the page numbers where each item can be found. The checklist can be downloaded from here: http://www.equator-network.org/reporting-guidelines/tripod-statement/

*We feel that the TRIPOD  checklist is not completely appropriate for our study because we use an infectious disease model rather than a statistical model designed to be used as a tool for prediction, thus several of the checklist items are not relevant. However, we would be happy to fill in this checklist if required.*

@@

*Thank you for your helpful comments on this paper.*

### Introduction

1.	Page 4, Line 21: BCG efficacy has been known with great heterogeneity and it affects the assessment of vaccine impact on TB burden. It is therefore commonly considered in decision-making for BCG vaccination policy. I suggest that the authors include the uncertainty of BCG efficacy in this analysis and compared the relative importance to the assumption of how TB burden declines over time.

*We agree with the reviewer that this is a potentially important issue. We have included some of this uncertainty by re-estimating incidence rates for both those BCG vaccinated and unvaccinated. These rates are then used to estimate vaccine efficacy. This approach makes use of the available data on the effiacy of the BCG vaccine in England and Wales. Including additional uncertainty observed globally would be inappropriate here as there is little evidence for this amongst the White UK born population. This does mean that the findings from this model reflect the impact in these countries only and cannot be directly generalised.*

@@

4.	Page 5, steps for defining the transmission model: Presenting equations of the model is certainly helpful for understanding the estimation methods. In addition to the descriptions of equation components, I encourage the authors to explain more about the rationales behind. For example, how to interpret ‘initial generation size (x)’? What is the purpose of relating step 1 to step 2? This kind of additional information will allow readers from a wide range of background to understand the model easily.

*Whilst this information was present in the original Sutherland et al. paper we agree that more detail was needed here. We have therefore added greater explanatory detail for each step (please see the marked up paper).*

5.	Page 5, last paragraph: Is the ‘distribution of secondary cases (N)’ estimated by the three age groups (of 15-19, 20-24, and 25-29 year old)? The power term ‘5-z’ of the decay percentage (d) needs additional explanations in the text.

*No. The distribution of secondary cases is not differentiated by age. In revising the manuscript, we identified an error in the original analysis, which we have corrected in our paper. This error means that this section of the model is no longer valid once the model has been updated and so it is only used to validate the original analysis. Based on this we have moved this section to the SI. However, we have expanded on the model explanation both here and in other sections.*

### 2019-10-11 README.Rmd

The original model greatly underestimated the impact of ending school-age vaccination compared to our updated, corrected, model. The model produced predictions with wide confidence intervals when parameter uncertainty was included. Estimates were most comparable to notification data when an annual decrease in TB incidence rates of 1.9% was used. Using this assumption, we estimate that 1600 (2.5,97.5% Quantiles (Q): 1300, 2000) vaccines would have been required to prevent a single notification in 2004.

@@

The previously published model was found to contain a methodological flaw which, once corrected, resulted in a large increase in the estimated impact of ending the BCG schools scheme. The original approach was also found to be spuriously precise once updated with parameter uncertainty. The impact of ending the BCG schools’ scheme was found to be greater than previously thought when notification data were used. Our results highlight the importance of independent evaluations of modelling evidence, including uncertainty, and evaluating multiple scenarios when forecasting the impact of changes in vaccination policy.

>>>>

The original model greatly underestimated the impact of ending school-age vaccination compared to our updated, corrected, model. The model produced predictions with wide confidence intervals when parameter uncertainty was included. Model estimates were most comparable to notification data when an annual decrease in TB incidence rates of 1.9% was used. Using this assumption, we estimate that 1600 (2.5,97.5% Quantiles (Q): 1300, 2000) vaccines would have been required to prevent a single notification in 2004.

@@

The previously published model was found to contain a methodological flaw which, once corrected, resulted in a large increase in the estimated impact of ending the BCG schools scheme. The original approach was also found to be spuriously precise once updated with parameter uncertainty. The impact of ending the BCG schools scheme was found to be greater than previously thought when notification data were used. Our results highlight the importance of independent evaluations of modelling evidence, including uncertainty, and evaluating multiple scenarios when forecasting the impact of changes in vaccination policy.
